---
title: 'Peer-graded Assignment: Course Project 1'
author: "Gabriela Manzella"
date: "28 February 2019"
output: html_document
---

####What is the mean total number of steps taken per day? 

The following chunk of code calculates the number of steps, the mean and the median per day using dplyr. 
It also plots a histogram for the number of steps per day. 

```{r}
library(dplyr)
library(rmarkdown)
library(knitr)
library(ggplot2)

actdata<-read.csv("./activity/activity.csv")
actdata$date<-as.Date(actdata$date)

mean_per_day<-actdata%>%
  na.omit()%>%
  group_by(date)%>%
  summarise(no_steps=sum(steps), mean_steps=mean(steps), 
            median_steps=median(steps))

histplot<-ggplot(data=mean_per_day, aes(x=no_steps))+
  geom_histogram(binwidth = 2000, colour="cadetblue4", fill="cadetblue2")+
  xlab("Number of steps per day")+
  ylab("Frequency")+
  scale_x_continuous(breaks=seq(0,25000, by=2500))+
  ggtitle("Histogram of number of steps per day")

ggsave("./figure/histplot.png", histplot)

```

```{r}
plot(histplot)
```

And the following tibble shows mean and median per day. 

```{r}
mean_per_day2<-mean_per_day[,c("date", "mean_steps", "median_steps")]

knitr::kable(mean_per_day2, format="html")
```

####What is the average daily activity pattern?

- How many NAs we have on the data set?

    + This was calculated by finding the number of NAs on the steps column:
        
        
```{r}
sum(is.na(actdata$steps))
```

- Fill in all the missing values.

    + This was done by filling in using the median of that day (which is 0 for all of them). 
  
```{r}

for (i in 1:length(actdata$steps)){
  if (is.na(actdata$steps[i])){
    actdata$steps[i]<-0
  }
} 

new_mean_per_day<-actdata%>%
  group_by(date)%>%
  summarise(no_steps=sum(steps), mean_steps=mean(steps), 
            median_steps=median(steps))

histplot2<-ggplot(data=new_mean_per_day, aes(x=no_steps))+
  geom_histogram(binwidth = 2000, colour="lightpink4", fill="lightpink2")+
  xlab("Number of steps per day")+
  ylab("Frequency")+
  scale_x_continuous(breaks=seq(0,25000, by=2500))+
  ggtitle("New histogram of number of steps per day by filling in the NAs")

ggsave("./figure/histplot2.png", histplot2)

plot(histplot2)

```
 
The plot looks about the same, but has a higher frequency of zeros, as expected. 

```{r}
new_mean_per_day2<-new_mean_per_day[,c("date", "mean_steps", "median_steps")]

knitr::kable(new_mean_per_day2, format="html")

```

####Are there differences in activity patterns between weekdays and weekends?

```{r}

actdata2<-actdata%>%
  mutate(day=weekdays(date), 
         daytype=ifelse(day=="Saturday" | day=="Sunday", "Weekend", "Weekday"))%>%
  group_by(interval, daytype)%>%
  summarise(average_steps=mean(steps))

plot3<-ggplot(data=actdata2, aes(x=interval, y=average_steps))+
  geom_line()+
  facet_wrap(~daytype)+
  ylab("Average number of steps per interval of time")+
  xlab("5 min interval")

ggsave("./figure/plot3.png", plot3)

plot(plot3)



```

From the plot above we can see that there are, in fact, differences between the activity patterns of the weekend and weekdays. The average is lower on a weekend, meaning they rest more. 